---
title: <span style="color:blue">Lab5</span>
output: 
    rmarkdown::html_vignette:
      toc: true

author: "Hoda, shwetha ,suhani"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{lab5}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE
)
```
The kolada database
-------------------

Within the kolada database there are two different types of data. 

* The municipal data 
* The organizational unit data

The municipal data contains KPIs for the Swedish kommuner and
landsting, there are about 4500 KPIs within the database. On
organizational unit level the data is less comprehensive, but there
are still quite a few KPIs on this level. Each KPI is currently
measured once a year, and the data may be divided by gender. I.e. for
each KPI there may be at most three values, based on gender being
female, male and a total.

source: https://github.com/Hypergene/kolada#kolada-api

<span style="color:blue">How to install</span>
=============

```{r , eval=FALSE}
devtools::install_github("shwva184/lab5")
runApp("inst", display.mode = "showcase")
```

<span style="color:blue">Required packages</span>
=================

```{r setup}
library(Lab5Assign)
library(dplyr)
require(shiny)
require(shinythemes)
```
### <span style="color:blue">How to Use</span>

to run the API interface :

```
{r }
runGitHub(repo = 'shwva184/lab5',username ='shwva184',subdir = 'inst', display.mode = "showcase")

```


#### Tab1 : KPI 

The Kolada API client provides access to standardized key performance indicators (KPI) concerning Swedish municipalities and organizational units.The follwing function is modified to give out KPI_Id,Tittle and description. For a proper query URL you need metadata such as id of a KPI(KPI Id) to retrieve the associated data.Tittle displays the name of the kpi. Description gives information needed to understand kpi.

Note the KPI id , which we will need to retrieve data from tab3

This functions fetches the above mention fields from URL : http://api.kolada.se/v2/kpi

```{r }
head(data.frame(Lab5Assign::getMetadata(entity ='kpi')),5) #Displaying only top 5 here.
```




### Tab2 : Municipality

Here we are retrieving information on ID associated with a particular municipality , names of all available muncipalities and municipality type is either L (for County Council (swedish: Landsting)) or K for municipality (swedish: Kommun).

Note the KPI id , which we will need to retrieve data from tab3

This functions fetches the above mention fields from URL : http://api.kolada.se/v2/municipality


```{r }
head(data.frame(Lab5Assign::getMetadata(entity ='municipality')),5)#Displaying only top 5 here.
```

### Tab3 : KPI and Municipality
The municipal data contains KPIs for the Swedish kommuner and landsting, there are about 4500 KPIs within the database. Each KPI is currently measured once a year, and the data may be divided by gender. I.e. for each KPI there may be at most three values, based on gender being female, male and a total.

The structure of the data is::
```
"KPI ID","Municipality ID","Year","Count(No. of members)","Gender", "Status","Value"
```
KPI_ID :  ID associated with KPI 
Municipality_ID: ID associated with Municipality
Year: the year KPI has been measured.
Count: number of members associated with the selected Municipality and KPI
Gender: Man, Kvinna and Total
Status column: shows if the average is available.
Value column : contains the unweighted average.(An unweighted average describes the average for Sweden's municipalities / regions - what it looks like in an average municipality / region. It is calculated by summing the value of all municipalities or regions and dividing by 290 and 21, respectively. If one or more municipalities / regions do not have data for the current key figure, the value is divided by the number that has values. Example: An unweighted average value for all municipalities regarding Merit value in year 9 will describe the value for an average municipality.)


To retrieve data for a specific kpi(ie. N02924) in a municipality(ie. 1440)
URL : http://api.kolada.se/v2/data/kpi/N02924/municipality/1440

```{r }
head(fetch_given_kpiandmuncipality_id("N02924",1440),10)#Displaying only top 10 here.
```
