---
title: "lab5"
output: rmarkdown::html_vignette
author: "Hoda, shwetha ,suhani"
vignette: >
  %\VignetteIndexEntry{lab5}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
The kolada database
-------------------

Within the kolada database there are two different types of data. 

* The municipal data 
* The organizational unit data

The municipal data contains KPIs for the Swedish kommuner and
landsting, there are about 4500 KPIs within the database. On
organizational unit level the data is less comprehensive, but there
are still quite a few KPIs on this level. Each KPI is currently
measured once a year, and the data may be divided by gender. I.e. for
each KPI there may be at most three values, based on gender being
female, male and a total.

The general structure of the data is::

    KPI, municipal or organizational unit, year, gender: data
source: https://github.com/Hypergene/kolada#kolada-api

# To install:

```{r , eval=FALSE}
devtools::install_github("shwva184/lab5")

```
$~$

# How to Use:
```{r setup}
library(Lab5Assign)

```
```{r }
library(shiny)
library(shinythemes)
```
To retrieve information about all the municipality run:
```{r }
head(data.frame(Lab5Assign::getMetadata(entity ='municipality')))
```

To retrieve information about a specific municipality (ie.Helsingborg ) run:

```{r }
Lab5Assign::getMetadata("municipality","Helsingborg")
```
To retrieve information about all the kpi in dataset 
run:
```{r 'kpi'}
head(data.frame(Lab5Assign::getMetadata(entity ='kpi')))

```

To retrieve information about a specific kpi(ie. )
